---
title: "ESW Data Analysis and Visualization"
author: "Jeho Park"
date: "4/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Raw Energy and Temperature Data

Pomona College's Studio Art building energy data.

The CSV file, Pomona_StudioArt_EnergyDemand_1yr.csv, contains raw energy and temperature data for different components

- Studio Art = total energy demand

- PV = solar panels

- Chiller = Machine for air conditioning

- MDB-AHU = air handling units

- LDB = light distribution board 

- H1 = other room lighting

In this workshop, we will 

- Import csv data file in to a data frame.

- Examine the raw data.

- Subset the data into different data frame objects using tidyverse packages (Data Wrangling).

- Plot demand and temperature in one graph using ggplot2 package.

## Get Ready (w/ RStudio Project)

(1) Download the csv file into your desktop or 
(2) Open RStudio
(3) Create a new project (File > New Project)
(4) Move the csv file into your project folder

## Import CSV data
(1) Click on the csv file
(2) Select Import Dataset...
(3) Check out the options and rename the data set `df`

The data set you just imported is now data frame object in your environment. (Check out the Global Environment)

Columns are variables and rows are observations. Let's check out the names of the columns of the data frame:
```{r class}
names(df)
```

## Data Preparation

### Name Change
Let's simplify the names to something meaningful yet shorter than current ones.
```{r names, eval=FALSE}
header <- c("Timestamp","SA.demand","PV.demand","CH.demand","MDB.demand","LDB.demand","H1.demand","LDB.temp","CH.temp","H1.temp","MDB.temp","PV.temp","SA.temp")
names(df) <- header
names(df)
```

### Subset the data set
Let's create a new data frame that only contains Studio Art's demand and temperature columns along with Timestamp column. There are different ways to subset data frame. The following shows two simple ways.

```{r sa_subsetting, eval=FALSE}
sa <- subset(df, select = c("Timestamp","SA.demand","SA.temp"))
sa1 <- df[,c(1,2,13)]
# Are these identical?
identical(sa, sa1)
```

Now let's do this subsetting for all the other sensors (devices) data. Remember we have six sensors (devices): `SA`, `PV`, `CH`, `MDB`, `LDB`, and `H1`.

```{r all_subsetting, eval=FALSE}
pv <- subset(df, select = c("Timestamp","PV.demand","PV.temp"))
ch <- subset(df, select = c("Timestamp","CH.demand","CH.temp"))
mdb <- subset(df, select = c("Timestamp","MDB.demand","MDB.temp"))
ldb <- subset(df, select = c("Timestamp","LDB.demand","LDB.temp"))
h1 <- subset(df, select = c("Timestamp","H1.demand","H1.temp"))
```

Now we want to see simple summary stats of these data.
```{r summary_stats, eval=FALSE}
summary(pv)
```
Something doesn't seem to right... Look at the demand class. It is always recommended to check the class of converted/imported data frame's columns.

How can we change the character objects to numeric objects?

```{r numeric, eval=FALSE}
sa$SA.demand <- as.numeric(sa$SA.demand)
pv$PV.demand <- as.numeric(pv$PV.demand)
ch$CH.demand <- as.numeric(ch$CH.demand)
mdb$MDB.demand <- as.numeric(mdb$MDB.demand)
ldb$LDB.demand <- as.numeric(ldb$LDB.demand)
h1$H1.demand <- as.numeric(h1$H1.demand)
```

## Data Visualization with Simple Plotting

Let's try plotting our `sa` data frame. 

```{r sa_plot, eval=FALSE}
library(ggplot2) # requires for qplot
# install.packages("ggplot2")
qplot(Timestamp, SA.demand, data=sa, geom="line") 
```

### Further Subsetting 

I see two problems here: NAs and too many data points. There are missing values in the data set. `SA.demand` and the plot shows the one full year of data, so there are too many data points. I can subset around a month, for example, August:

```{r sa_aug, eval=FALSE}
sa.aug <- subset(sa, 
                 format(`Timestamp`, "%Y-%m-%d") >= "2017-08-01" &
                 format(`Timestamp`, "%Y-%m-%d") < "2017-09-01")

qplot(Timestamp, SA.demand, data=sa.aug, geom="line") 
```

For more information about subsetting data frame, see https://www.r-bloggers.com/5-ways-to-subset-a-data-frame-in-r/ for example.

## Tidy Data

See https://ramnathv.github.io/pycon2014-r/explore/tidy.html

> What is Tidy Data?
> A dataset is said to be tidy if it satisfies the following conditions:
> 
> - observations are in rows
>
> - variables are in columns
>
> - contained in a single dataset.
>
> Tidy data makes it easy to carry out data analysis.
>

Look at this example:

```{r sa_tidy, eval=FALSE}
library("reshape2") # need for melt

sa_tidy <- melt(sa, id = "Timestamp", 
                variable.name = "type", 
                value.name = "value", 
                na.rm = TRUE)
head(sa_tidy)
```

Let's subset August 2017 data using the tidy data:

```{r sa_tidy_subset, eval=FALSE}
# subet sa_tidy into August data
sa_tidy.aug <- subset(sa_tidy, format(Timestamp, "%m") == "08", na.rm = TRUE)

# subset sa into demand and temp
sa_tidy.aug.demand <- subset(sa_tidy.aug, type=="SA.demand")
sa_tidy.aug.temp <- subset(sa_tidy.aug, type=="SA.temp")

# quick plot for SA Demand and Temp
qplot(Timestamp, value, data=sa_tidy.aug.demand, geom="line")
qplot(Timestamp, value, data=sa_tidy.aug.temp, geom="line")

# Use ggplot to display two values corresponding to different types
ggplot(data=sa_tidy.aug,
       aes(x=Timestamp, y=value, colour=type)) +
  geom_line()

```

## Export data frame as a CSV file
```{r data_export, eval=FALSE}
write.csv(sa_tidy, file = "sa_tidy.csv")
```

## Adding secondary axis
```{r secondary_axis, eval=FALSE}
p <- ggplot(data=sa_tidy.aug,
       aes(x=Timestamp, y=value, colour=type)) +
  geom_line()

p + scale_y_continuous("MSCI Value", sec.axis = sec_axis(~ . /300, name = "Interest Rate")
p
```
